SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v20]

### 2015-09-10 paper.tex
added Ken's comments ...v20

\usepackage{tabularx}
\usepackage{tikz,hyperref,graphicx,units}
\usepackage{subfigure}
\usepackage{benktools}

@@

\newboolean{include-notes}
\setboolean{include-notes}{true}
\newcommand{\adnote}[1]{\ifthenelse{ \boolean{include-notes}}%
 {\textcolor{blue}{\textbf{AD: #1}}}{}}

@@

\title{SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v19]}

@@

\begin{abstract}
Online algorithms for robot model-free robot learning from demonstration (such as DAgger) can learn policies for problems where the system
dynamics and the cost function are unknown. However, during learning they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states when iteratively executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by enabling the robot to \emph{only request supervision for risky states}. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936D visual HOG feature state representation, 2) 4DOF robot arm push-grasping in clutter (simulation with Box2D) with a 22D state representation  of joint angles and the pose of each object, and 3) surgical needle insertion with a 16D state represented spatial transform.  Results suggest that SHIV can reduce the number of queries up to 70$\%$.

%In traditional imitation learning, a human supervisor provides a few demonstrations to the robot, which then \emph{offline} learns a policy to extrapolate these examples to new situations. One challenge is that during execution of the learned policy, a small but inevitable error will amplify, taking the robot further and further away from the regions of state space for which it has seen examples.

%\emph{Online} imitation learning algorithms such as DAgger address this challenge by iteratively executing a trained policy, and asking the supervisor for additional examples along the states visited by that policy. Such algorithms perform well in new situations, but rely much more heavily on extensive human supervision.

%Our goal is to reduce the amount of human supervision required in imitation learning, without sacrificing performance. \emph{Our key insight is that the robot should ask for supervision only when its current trained policy has high risk.}

%We introduce a method for online imitation learning that evaluates risk based on novelty of a state, as well as prior difficulty in labeling a nearby states. Our algorithm, SHIV (SVM-Based Reduction in Human Intervention), follows the trained policy when evaluated risk is low, and asks for human supervision otherwise.

%A trained policy can have a high risk of predicting the wrong action for a state for two reasons: 1) the state is too far away from the training states, and 2) the state is in an area in which the learner has historically struggled. We introduce a method for iterative imitation learning that estimates regions of high risk based on these two criteria, and limits queries to the supervisor to only high risk states. Our method instantiates stream based active learning, with a query selection strategy that generalizes novelty detection to querying in areas that are inherently difficult to correctly label.

% \adnote{Old abstract:}

% For robot control problems where the reward function  and the dynamics of the environment are not specified, the
% learning from demonstrations approach provides an algorithmic paradigm for learning a policy from supervisor-provided
% demonstrations. Dataset Aggregation (DAgger) in particular yields a state of the art method in this context which treats
% the problem of determining a policy as an iterative supervised learning problem which alternates between executing a
% trained policy and acquiring new supervisor control inputs for roll-outs of a current best policy estimator. In this
% work, we propose a novel algorithm, \acro~which is based on DAgger and which estimates the riskiness of states
% encountered by the policy roll-outs by means of an approximate quantile super-levelset estimate of the distribution of
% states used for training the current policy. Since learning a quantile function for high-dimensional input data is
% challenging, we instead modify an SVM-based approach to quantile estimation that employs regularization to solve this
% problem approximately. Our algorithm considers roll-outs with early termination as soon as a state outside a
% user-defined quantile super-levelset is encountered. Our approach furthermore reduces the amount of neccessary
% supervisor demonstrations by only collecting supervisor control inputs at states outside the current quantile estimate.
% In experiments involving a simulated race track and in a Super Mario Atari game benchmark considered by the authors of
% DAgger, our approach results in similar performance levels as DAgger while reducing the amount of required supervisor
% control inputs by more than {\color{blue}[-- TODO --]} percent.
 \end{abstract}

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment (Fig. \ref{fig:teaser}), from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to neither the cost function that it should optimize, nor the dynamics model. The former happens when it is difficult for a human to specify how to trade-off various aspects that matter, like when the robot is reaching for a goal object while not pushing objects off a table(Fig. \ref{fig:teaser}(b).  The latter happens when either the system or the interaction with the world is difficult to characterize with the given observations, like in Fig. \ref{fig:teaser}(a) the robot must stay on the track but only has access to visual information from a bird's eye view.

Rather than learning the cost function and the dynamics model, and then using optimization to produce a policy for the task, the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}.

@@

In the offline learning from demonstration setting, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate and push the robot away from the region of the state space where it has seen examples. The learned policy might not be helpful in unfamiliar states, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. The number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online learning from demonstration approaches such as DAgger address this issue by iteratively gathering more examples from the supervisor in states the robot is likely to visit~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. DAgger learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy and the supervisor provides demonstrations for all states the robot visits. The new examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, the number of errors scales only linearly with the time horizon of the task.

To achieve better performance guarantees, online algorithms can significantly increase the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading robot performance.

Our key insight, to reduce burden, is that the robot only needs to request demonstrations for risky states. Risk arises because:  (1) states are different than the states previously encountered and used to train the current policy, i.e., they can be considered novel or outliers~\cite{hodge2004survey}; or (2) states are in a region that has been misclassified, in previous training data.
%\adnote{Ken's version below does not seem to represent what we actually do.}
%the action associated with a state under the current policy yield a resulting state that is inconsistent with the expected next state. [how do we measure 2? - Ken]

We contribute an algorithm for online learning from demonstration that can actively decide whether it is necessary to ask the supervisor to label the new states that the robot encounters. Our algorithm, SHIV  (Svm-based reduction in Human InterVention), reduces supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to the non-stationary state distributions the robot encounters as it iteratively executes learned policies. We build on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

\mlnote{Ken didn't want to repeat the abstract here, so we removed this section}
\adnote{we need results here; both the highlights of the analysis we do (so that people don;t think we are blindly comparing it to dagger without asking if all these changes are necessary), as well as the main point, that shiv is better for the same budget, and refer to fig 1}

@@

@@

\mlnote{Ken is a big fan of  discussing the assumptions and model choices without notation before everything else, so a reader can quickly determine if this method suits there problem}

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as markovian, stochastic, and stationary. Stationary meaning  given a state and a control the probability of the next state doesn't change over time. We model the initial state as sampled from a distribution over the state space.

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide a control signal label. We assume the supervisor the supervisor can be noisy and imperfect, but also that the robot can at best reach the same level of performance of the supervisor after learning.

\noindent\textbf{Policies and State Densities.}
We denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics, as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ from time $0$ to time $t-1$. This density can be computed by marginalization: $p(\bx_t|\theta) =
\int_{\bx_{t-1}}...\int_{\bx_1} p((\bx_t,...,\bx_1)|\theta) d\bx_{t-1}...d\bx_1$. Following \cite{ross2010reduction}, we can compute
the average density on states \emph{for any timepoint} by 
\begin{equation}
p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)
\label{eq:density}
\end{equation}

While we do not assume analytic knowledge of the distributions corresponding to: $p(\bx_{t+1}|\bx_t,\bu_t)$, $p(\bx_0)$, $p(\bx_t|
\theta)$ or $p(\bx|\theta)$, we assume that we have a stochastic real robot or a simulator such that for any state
$\bx_t$ and control $\bu_t$, we can sample the $\bx_{t+1}$ from the density $p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$. 
Therefore, when 'rolling out' trajectories under a policy
$\pi_{\theta}$, we utilize the robot or a simulator to sample the resulting stochastic trajectories rather than
estimating $p(\bx|\theta)$ itself.

However in our learning from demonstration setup, the robot does not have access to the cost function itself. Instead, it only has access to 
a supervisor that we assume uses some $\tilde{\pi}$ to minimize $C(\hat{\tau})$, to an acceptable level. \mlnote{the last statement is intended to mean implementors of this algorithm have agreed that the supervisor may not be optimal, but has a level of performance that is worth automating} We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training dataset $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a policy that minimizes the surrgoate loss between relative to the supervisor's policy.

Following \cite{ross2010reduction}, our objective then consists of determining a policy $\pi_{\theta}$ minimizing the expected surrogate loss, where the expectation is taken over the distribution of states induced by the policy across any time point in the horizon:

 If the robot could learn the policy  perfectly, this state density would match the one encountered in the user examples. But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to \cite{pomerleau1989alvinn}.Thus, motivating iterative algorithms like DAgger and now SHIV, which iterate between learning a policy and then the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.

% The optimization problem in Eq. \ref{eq:LFD_obj} is  typically non-convex. It can also be difficult to solve, since we only have access to samples from the supervisor's policy $\tilde{\pi}$ arising from the demonstration trajectories $\mathcal{D}$. To address these issues Stephane and Ross proposed an iterative method known as DAgger \cite{ross2010reduction}.

@@

\subsection{Step 1.}
The first step of any iteration $k$ is to compute a $\theta_k$ that minimizes surrogate loss on the current dataset $\mathcal{D}_k=\{(x_i,u_i)|i\in\{1,\ldots,M\}\}$ of demonstrated state-control pairs (initially just the set $\mathcal{D}$ of initial trajectory demonstrations):

@@

This problem is a supervised learning problem, solvable by estimators like a support vector machine or a neural net \footnote{To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
can be considered as present in the policy's output.  A regularization technique in the optimization is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.}.  Finding the right function representation and optimization method for a given problem is still an open question in machine learning. However a large number of advances have been made that make the use of these techniques very feasible \cite{scholkopf2002learning} \mlnote{The point was to acknowledge to traditional machine learning people that the problem is still hard and not solved. Ken didn't want to offend them}\adnote{is this really needed? what does it add? it basically says figuring out what method to use is hard; maybe say: performance will vary with the selection of a particular estimator \cite{}}

 \subsection{Step 2}
The second step starts in both SHIV and DAgger by rolling out the policy $\pi_{\theta_{k}}$ to sample states that are likely under $p(\bx|\theta_{k})$.

What happens next, however, differs. For every state visited, DAgger requests the supervisor to provide the appropriate control/label. Formally, for a given sampled trajectory  $\hat{\tau} = (\bx_0,\bu_0,...,\bx_T,\bu_T )$, the supervisor provides labels $\tilde{\bu}_t$, where $\tilde{\bu}_t \sim \pi(\bx_t) + \epsilon$ for $t\in \{0, \ldots, T\}$.
The states and labeled controls are then aggregated into the next data set of demonstrations $\mathcal{D}_{k+1}$:
$$D_{k+1}=\mathcal{D}_k \cup \{(\bx_t,\tilde{\bu_t})\|t\in\{0,\ldots,T\}\} $$
%\fpnote{terms such as aggregated etc would be much clearer if we just refer to a particular line in the algorithm}

Providing correct control inputs or ``labeling'' for all states encountered at each
iteration can impose a large burden on the supervisor.
Instead of asking the supervisor for labels at all visited states $\{\bx_0,..,\bx_T\}$, SHIV uses a measure of risk to actively decide whether a label is necessary. For every state encountered $\bx_t$, it applies a decision function $g_{\sigma}$, which we introduce in the next section, parametrized by a risk-sensitivity parameter $\sigma$ which captures how smooth we expect the policy function to be: smoother functions imply that training examples have larger support, and data nearby will be less risky.

SHIV only asks for supervision on states for which $g_{\sigma}(\bx) \neq 0$:

$$D_{k+1}=\mathcal{D}_k \cup \{(\bx_t,\tilde{\bu_t})\|t\in\{0,\ldots,T\},g(\bx_t)=-1\}$$

Steps 1 and 2 are repeated for $K$ iterations or until 
the cumulative surrogate is smaller than some predefined threshold\footnote{In the original DAgger the policy rolled out
was stochastically mixed with the supervisor, thus with probability $\beta$ it would either take the supervisor's action
or the robots. The use of this stochastically mix policy was for theoretical analysis. In practice, it is recommended
to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}.

 For a conservative $\sigma$, the SHIV performance is analogous to DAgger without asking the expert when it is absolutely not necessary. However, if $\sigma$ is too high, then SHIV will not collect new data that is needed and revert back to the performance of offline methods. We analyze the dependence on $\sigma$ in Sec. \ref{sec:car}, and our results suggest that SHIV is robust to this parameter selection.

\subsection{Measuring Risk}
% However, policy rollouts can in practice result in the robot entering dangerous states. For example, when DAgger was used on a
% quad-copter, the robot repeatedly crashed into trees before successfully learning to avoid them \cite{ross2013learning}.
We work under the assumption that a state can be risky for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which can cause our policy to mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance}, or 2) the surrogate loss, or training error, of the current policy at a state in $\mathcal{D}$ is high, so that the state does not model the supervisor's control inputs correctly.

To evaluate risk in high-dimensional state spaces (such as HOG features on images, like in our driving simulator, Fig. \ref{fig:teaser}(a)), we use
a modified version of the technique known as the One Class SVM that  approximately estimates a boundary of a user defined quantile of
a density representing the training data in $\mathcal{X}$ \cite{scholkopf2001estimating}.
%\fpnote{Refer to figure, replace "one way to do"  by "an approach to this problem is to" or similar - less informal}

>>>>

\usepackage{tabularx}
\usepackage{tikz,hyperref,graphicx,units}
\usepackage{subfigure}
\usepackage{benktools}
\usepackage{bbm}

@@

\newboolean{include-notes}
\setboolean{include-notes}{false}
\newcommand{\adnote}[1]{\ifthenelse{ \boolean{include-notes}}%
 {\textcolor{blue}{\textbf{AD: #1}}}{}}

@@

\title{SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v20]}

@@

\begin{abstract}
Online learning from demonstration algorithms, such as DAgger, can learn policies for problems where the system
dynamics and the cost function are unknown. However, during learning they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states when executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936 dimension visual HOG feature state representation, 2) 4DOF robot arm push-grasping in clutter simulation with Box2D with a 22 dimension state representation  of joint angles and the pose of each object, and 3) surgical needle insertion with a 16 dimension state represented spatial transform.  Results suggest that SHIV can learn equivalent policies with up to $70\%$ fewer queries.

 \end{abstract}

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former occurs when it is difficult to specify how to trade-off various aspects that matter, like when the robot is reaching for a goal object while not pushing objects off a table.  The latter occurs when either the system or the interaction with the world is difficult to characterize based on observations, as in Fig. \ref{fig:teaser}(a) the robot drives on the track but only has access to visual information from a bird's eye view.

Rather than explicitly learning the cost function, like in inverse reinforcement learning, and the dynamics model, and then using optimization to produce a policy for the task, the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}. Learning from demonstrations algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.

@@

In offline learning from demonstration, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online learning from demonstration addresses this issue by iteratively gathering more examples from the supervisor in states the robot encounters~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. One such algorithm, DAgger learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy and the supervisor provides demonstrations for all states the robot visits. The new state/control examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, it was proved under certain conditions, the number of errors scales only linearly with the time horizon of the task.

DAgger significantly increases the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading robot performance.

Our key insight is that the robot only needs to request demonstrations for risky states. Risk arises because:  (1) states are sufficiently different than the states previously encountered, i.e., they can be considered novel or outliers~\cite{hodge2004survey}; or (2) states are in a region that has been misclassified in previous training data.
%\adnote{Ken's version below does not seem to represent what we actually do.}
%the action associated with a state under the current policy yield a resulting state that is inconsistent with the expected next state. [how do we measure 2? - Ken]

We present an algorithm for online learning from demonstration that can actively decide whether it is necessary to ask the supervisor to label the new states that the robot encounters. SHIV  (Svm-based reduction in Human InterVention) reduces supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to the non-stationary state distributions the robot encounters as it iteratively executes learned policies. We build on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

@@

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as Markovian, stochastic, and stationary. Stationary dynamics meaning  given a state and a control the probability of the next state doesn't change over time. Note this is different then the non-stationary distribution over the states the robot encounters during learning. We model the initial state as sampled from a distribution over the state space.

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide a control signal label. We assume the supervisor can be noisy and imperfect. Noting that the robot cannot surpass the level of performance of the supervisor.

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics, as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ from time $0$ to time $t-1$.  Following \cite{ross2010reduction}, we can compute
the average density on states for any timepoint by $p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)$.

While we do not assume knowledge of the distributions corresponding to: $p(\bx_{t+1}|\bx_t,\bu_t)$, $p(\bx_0)$, $p(\bx_t|
\theta)$ or $p(\bx|\theta)$, we assume that we have a stochastic real robot or a simulator such that for any state
$\bx_t$ and control $\bu_t$, we can sample the $\bx_{t+1}$ from the density $p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$. 
Therefore, when 'rolling out' trajectories under a policy
$\pi_{\theta}$, we utilize the robot or a simulator to sample the resulting stochastic trajectories rather than
estimating $p(\bx|\theta)$ itself.

However in this case, we do not have access to the cost function itself. Instead, we only have access to 
a supervisor that can achieve a desire level of performance on the task.  We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training dataset $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a policy that minimizes the surrogate loss relative to the supervisor's policy.

Following \cite{ross2010reduction}, our objective is to find a policy $\pi_{\theta}$ minimizing the expected surrogate loss, where the expectation is taken over the distribution of states induced by the policy across any time point in the horizon:

 If the robot could learn the policy  perfectly, this state density would match the one encountered in user examples. But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to \cite{pomerleau1989alvinn}.This motivates iterative algorithms like DAgger, which iterate between learning a policy and the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.

\section{Risky and Safe States}

@@

Providing correct control inputs or ``labeling'' for all states encountered at each
iteration can impose a large burden on the supervisor.
Instead of asking the supervisor for labels at all visited states, SHIV uses a measure of risk to actively decide whether a label is necessary.

We define a state as "risky" for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which can cause the current policy to mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance}, or 2) the surrogate loss, or training error, of the current policy at the state  is high, so that the state is unlikely to model the supervisor's control inputs correctly. States that are not classified as "risky" are deemed "safe".

The amount of data needed to estimate the density, scales exponentially in the dimension of the state space \cite{nadaraya1964estimating}. 
Thus, to evaluate risk in high-dimensional state spaces , such as the HOG features in our driving simulator, Fig. \ref{fig:teaser}(a), we use
a modified version of the technique known as the One Class SVM that  estimates a regularized boundary of a user defined quantile on the training data in $\mathcal{X}$ \cite{scholkopf2001estimating}.
%\fpnote{Refer to figure, replace "one way to do"  by "an approach to this problem is to" or similar - less informal}

%\subsection{Estimation of Quantile Level Sets}\label{sec:level}
We consider the problem of estimating the quantile level-sets of a distribution $P$ on a set $\mathcal{X}$ by means of a finite set of
independent and identically distributed samples $\mathbf{x}_1,...,\mathbf{x}_n\in \mathcal{X}$.
In most general terms, the quantile function for $P$ and subject to a class of measurable subsets $\mathcal{G}$ of $\mathcal{X}$ is
defined by
\vspace{-2ex}
\begin{align}\label{eq:quantile}
U(\gamma) = \mbox{inf} \lbrace \lambda(G):P(G) \geq \gamma, G \in \mathcal{G} \rbrace \: 0<\gamma \leq 1
\end{align} 
$\lambda:\mathcal{G}\to \mathbb{R}$ above denotes a volume measure.
Suppose furthermore that $G:[0,1]\to \mathcal{G}$ assigns a set $G(\gamma) \in \mathcal{G}$ that attains the infinum
measure (i.e. volume) for each $\gamma\in [0,1]$ (this set is in general not necessarily unique). 
$G(\gamma)$ denotes a set of minimum measure $G \in \mathcal{G}$ with $P(G(\gamma))\ge \gamma$. Note in particular that $G(1)$ is the support of the density $p$ corresponding to $P$, if $p$ exists.

To handle distributions defined on high-dimensional spaces $\mathcal{X}$, work by Scholk{\"o}pf et al. represents the class $\mathcal{G}$ via a kernel $k$ as the set of half-spaces in the support vector (SV) feature space \cite{scholkopf2001estimating}. 
By minimizing a support vector regularizer controlling the smoothness of the estimated level set function this work
derives an approximation of the quantile function described in Eq. \ref{eq:quantile}.
This approach can be thought of as employing $\lambda(G) = ||w||^2$, where $G_w = \lbrace x: f_w(x) \geq \rho \rbrace$,
$f_w(\mathbf{x}) = \sum_i w_i k(\mathbf{x}_i, \mathbf{x})$
and $(w,\rho)$ denote the weight vector and offset parameterizing a hyperplane in the feature space associated with a
Gaussian kernel $k(\bx_0,\bx_1) = e^{-||\bx_0 - \bx_1||^2/2\sigma^2}$.

Here, the parameter $\nu$ controls the penalty or `slack term' and is equivalent to $\gamma$ \cite{vert2006consistency}
in the quantile definition, Eq. \ref{eq:quantile}, as the number of samples increases. The decision
function, determining point membership in the approximate quantile levelset is given by $g(\bx) = \mbox{sgn}(\langle w,\Phi(x) \rangle-\rho)$. Here, for $x\in \mathcal{X}$, $g(x)=0$ if $x$ lies on the quantile levelset,
$g(x) = 1$ if $x$ is strictly in the interior of the quantile super-levelset and $g(x) = -1$ 
if $x$ lies strictly in the quantile sub-levelset.

The dual form of the optimization yields a Quadratic Program 
that has worst case computational complexity of $O(n^3)$. However, Sch{\"o}lkopf et al. developed an improved optimization method that has empirically been shown to be as low as $O(n^2)$ \cite{scholkopf2001estimating}. In the dual, the decision function is given by $g(\bx) = \mbox{sgn}(\sum^N_{i=1}\alpha_i k(\bx_i,\bx)-\rho)$ where $\alpha_i$ corresponds to the dual variables.  The novelty detection method can be visualized in Fig. \ref{fig:support_example}(c). 
However, even when sufficient data is available, the associated control inputs may be inconsistent or noisy and a resulting policy
optimizing Eq. \ref{eq:super_objj} may still incur a large surrogate loss. To account for this, we propose a
modification to the One Class SVM:

\vspace{-2ex}
\begin{align}
y_i = \left\{
     \begin{array}{lr}
         1 & : l(\pi_{\theta}(\bx_i),\bu_i)\le \varepsilon\\
         -1 & : l(\pi_{\theta}(\bx_i),\bu_i)>\varepsilon
     \end{array}
   \right.
\end{align}

We divide up our data in to two sets those correctly classified: $\mathcal{D}_{s}=\{\lbrace \bx_i,\bu_i \rbrace \in \mathcal{D}_{k}, y=1\}$ and those states incorrectly classified: $\mathcal{D}_{r}=\{\lbrace \bx_i,\bu_i \rbrace \in \mathcal{D}_{k}, y=-1\}$
A separate One-Class SVM is then trained on each set of states, ($D_{s}$ and $D_{r}$) and providing measures of the level sets, $g_{s}$ and $g_{r}$. Specified by parameters $(\nu,\sigma)$ and $(\nu_r,\sigma_r)$, respectively.

\vspace{-2ex}
\begin{align}\label{eq:decision_func}
g_{\sigma}(\bx) = g_{r} (\bx)+ g_{s}(\bx)
\end{align}
points are deemed risky if $g_{\sigma}(\bx) \neq 0$.  Practically, this modification corresponds to
`carving out holes' in the estimated quantile super-levelset such that neighborhoods around states with $y_i=-1$ are
excluded from the super-levelset. An illustration of this can be seen in Fig. \ref{fig:support_example}(d).

The decision function parametrization consists of the kernel bandwidth $\sigma$ in $g_s$. We treat $\sigma$ as a "risk sensitivity" parameter. For two reasons: 1)The expected number of examples, after a policy roll out, the supervisor can be asked is  $T*\int_\bx \mathbbm{1}(g_{\sigma}(\bx) == 0) p(\bx|\theta)d\bx$. Thus, smaller $\sigma$ corresponds to asking for more examples. 2) A relation exists between how smooth the supervisor's poilcy, $\tilde{\pi}$ and how many examples are needed to learn it. Thus, a large $\sigma$ can be dangerous for policies with sharp variation because it will treat points as safe that are really risky.

@@

\subsection{Step 1}
The first step of any iteration $k$ is to compute a $\theta_k$ that minimizes surrogate loss on the current dataset $\mathcal{D}_k=\{(x_i,u_i)|i\in\{1,\ldots,M\}\}$ of demonstrated state-control pairs (initially just the set $\mathcal{D}$ of initial trajectory demonstrations):

@@

This sub-problem is a supervised learning problem, solvable by estimators like a support vector machine or a neural net \footnote{To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
can be considered as present in the policy's output.  A regularization technique in the optimization is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.}. Performance can vary though with the selection of a the estimator \cite{scholkopf2002learning}

 \subsection{Step 2}
The second step SHIV and DAgger rolls out their policies, $\pi_{\theta_{k}}$, to sample states that are likely under $p(\bx|\theta_{k})$.

What happens next, however, differs between SHIV and DAgger. For every state visited, DAgger requests the supervisor to provide the appropriate control/label. Formally, for a given sampled trajectory  $\hat{\tau} = (\bx_0,\bu_0,...,\bx_T,\bu_T )$, the supervisor provides labels $\tilde{\bu}_t$, where $\tilde{\bu}_t \sim \tilde{\pi}(\bx_t) + \epsilon$, where $\epsilon$ is a small zero mean noise term, for $t\in \{0, \ldots, T\}$.
The states and labeled controls are then aggregated into the next data set of demonstrations $\mathcal{D}_{k+1}$:
$$D_{k+1}=\mathcal{D}_k \cup \{(\bx_t,\tilde{\bu_t})\|t\in\{0,\ldots,T\}\} $$
%\fpnote{terms such as aggregated etc would be much clearer if we just refer to a particular line in the algorithm}

SHIV only asks for supervision on states for which are risky, or $g_{\sigma}(\bx) \neq 0$: 
$$D_{k+1}=\mathcal{D}_k \cup \{(\bx_t,\tilde{\bu_t})\|t\in\{0,\ldots,T\},g(\bx_t)=-1\}$$
Steps 1 and 2 are repeated for $K$ iterations or until 
the robot has achieved sufficient performance on the task\footnote{In the original DAgger the policy rolled out
was stochastically mixed with the supervisor, thus with probability $\beta$ it would either take the supervisor's action
or the robots. The use of this stochastically mix policy was for theoretical analysis. In practice, it is recommended
to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}.
